### 2025-07-11 pca.qmd
20250711 - PCA interpretation/labelling the components

@@

@@

@@

@@

@@

@@

>>>>

The variables in component 1 appear to reflect Quarterback usage: completions, incompletions, passing attempts, passing yards, passing touchdowns, interceptions thrown, fumbles, times sacked, sack yards lost ("reversed"—i.e., negatively associated with the component), sack fumbles, sack fumbles lost, passing air yards (total horizontal distance the ball travels on all pass attempts), passing yards after the catch, and first downs gained by passing.
Quarterbacks who tend to throw more tend to have higher levels on those variables.
Thus, we label component 1 as "Usage", which reflects total Quarterback involvement, regardless of efficiency or outcome.

The variables in component 2 appear to reflect Quarterback aggressiveness: passing air yards, passing air conversion ratio (reversed; ratio of passing yards to passing air yards), average air yards on completed passes, average air yards on all attempted passes, aggressiveness (percentage of passing attempts thrown into tight windows, where there is a defender within one yard or less of the receiver at the time of the completion or incompletion), expected completion percentage (reversed; based on air distance, receiver separation, Quarterback/Wide Receiver movement, pass location, whether there was pressure on the Quarterback when throwing, the throw angle and trajectory, receiver and defender positioning at the catch point, and defensive coverage scheme), average amount of air yards ahead of or behind the first down marker on passing attempts, average air distance (the true three-dimensional distance the ball travels in the air), maximum air distance, and maximum air distance on completed passes.
Quarterbacks who throw the ball farther and into tighter windows tend to have higher values on those variables.
Thus, we label component 2 as "Aggressiveness", which reflects throwing longer, more difficult passes with a tight window.

The variables in component 3 appear to reflect Quarterback performance: passing touchdowns, passing expected points added, passing completion percentage above expectation, passes completed of 40 yards or more, pass completion percentage, air yards differential (intended air yards $-$ completed air yards; attempting deeper passes than he on average completes); maximum completed air distance, average time to throw, average completed air yards, average intended air yards, and passer rating.
Quarterbacks who perform better tend to have higher values on those variables.
Thus, we label component 3 as "Performance".

@@

Here are the players and weeks that showed the highest levels of Quarterback "Usage":

@@

Here are the players and weeks that showed the lowest levels of Quarterback "Usage":

@@

Here are the players and weeks that showed the highest levels of Quarterback "Aggressiveness":

@@

Here are the players and weeks that showed the lowest levels of Quarterback "Aggressiveness":

@@

Here are the players and weeks that showed the highest levels of Quarterback "Performance":

@@

Here are the players and weeks that showed the lowest levels of Quarterback "Performance":
